Random Forests", OAs

### 2023-11-23 src/oa_RF.Rmd
CAST

# Partially complete panel as not all OAs have a website in at least ine year
df <- df %>% filter(!is.na(oa11cd)) %>% 
  complete(oa11cd, year) %>% 
  replace(is.na(.), 0)

@@

```{r}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

@@

# Build OA for GB
oa.gb <- rbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

@@

```{r}
dist <- distm(cbind(oa.gb.c$Y, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

@@

@@

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol", 
#                         "Manchester", "Sheffield", "Leeds", "Edinburgh", 
#                         "Leicester", "nearest.city.oa11cd")) %>% 
#   write_csv(file = out.path)
```

>>>>

# Partially complete panel as not all OAs have a website in at least one year
df <- df %>% filter(!is.na(oa11cd)) %>% 
  complete(oa11cd, year) %>% 
  replace(is.na(.), 0)

@@

```{r eval=F}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

@@

# NI
path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
  rename(id= SA2011,
         area = Hectares)

# Build OA for GB
oa.gb <- rbind(oa.ew, oa.sc, oa.ni)
rm(oa.ew, oa.sc, oa.ni)

@@

```{r}
dist <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

@@

# in.path <- paste0(path, "/data/temp/df_for_oa_rf_old.csv")
# df.compare <- read_csv(in.path)

@@

df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
                        "Leicester", "nearest.city.oa11cd"))

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)
```

## RF with CAST

```{r}
df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
  mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

df <- df %>% filter(year > 1996)

#df <- df %>% dplyr::select(-n.lag)
```

### unregister_dopar function

This is helpful when the parallel loop collapses.

```{r}
unregister_dopar <- function() {
  env <- foreach:::.foreachGlobals
  rm(list=ls(name=env), pos=env)
}
unregister_dopar()
```

### train the model in all data

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "oa11cd",
                                timevar = "year", 
                                k = 10)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

# detectCores()
# cl <- makePSOCKcluster(2)
# registerDoParallel(cl)

start.time <- Sys.time()
model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

# stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

print(model.all)

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.slag", "spatial lag of web density",
                                               ifelse(rowname == "n.lag", "temporal lag of web density",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag  of web density",
                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/varimp_OA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)

# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
#                                                         "year")], 
#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
#                     trControl = tc,
#                     method = "ranger",
#                     #na.action = na.omit,
#                     #importance = "impurity",
#                     metric="Rsquared",
#                     verbose=FALSE)
# ffsmodel_LLO
# ffsmodel_LLO$selectedvars
# plot_ffs(ffsmodel_LLO)
```

### train the model in a loop per region

```{r}

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

# train in every region
start.time <- Sys.time()

regions <- df %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
                                spacevar = "oa11cd",
                                timevar = "year", 
                                k = 10)

  model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df %>% filter(RGN21CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all)
}

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken
```

### Test on all other regions

```{r}
# train on one region, test on all the rest

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") + #y: test.region, x: train.region
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
